Malacology: A Programmable Storage System

### 2016-10-18 paper/paper.tex
Adds discussion on SDS and Active Storage
Fixes #93

%These three trends put evolutionary pressure on storage systems and raise the
%question of whether 
The evolutionary pressure placed on storage systems by these trends raises the
question of whether
there are principles that storage systems designers can follow
to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.
% that are important for
%users to continue to entrust their data to the storage system.
In this paper we investigate an approach that focuses on 
%generalizing 
%and
%composing 
identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. 
%We will refer to this programmability as
%\emph{programmable storage}, which differs from \emph{active storage} (the
%injection and execution of arbitrary code in a storage system) 
%and
%\emph{software-defined storage} (the control of thin-provisioning of storage).

{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
re-use and extension of existing storage abstractions provided by the underlying software stack,
to enable the creation of new services via composition.
A programmable storage system can be realized by exposing existing functionality (such as 
metadata services and synchronization and monitoring capabilities) as interfaces that can be ``glued together''
in a variety of ways
using a high-level language. In contrast to programmable storage, 
\emph{active storage} differs through its focus on the injection and
execution of code within a storage system or storage device.

@@

Programmable storage systems can be viewed as an infrastructure for creating
abstractions to better separate policies from mechanisms. This idea is not new.
Software-defined networks (SDNs) create such an abstraction by separating the
control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
notion of control/data separation was also applied in software-defined storage
(SDS)~\cite{arpaci:sosp01,thereska:sosp13,stefanovici:fast16}. Similarly,
IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
filtering in OpenStack Swift.

>>>>

%These three trends put evolutionary pressure on storage systems and raise the
%question of whether 
The evolutionary pressure placed on storage systems by these trends raises the
question of whether
there are principles that storage systems designers can follow
to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.
% that are important for
%users to continue to entrust their data to the storage system.
In this paper we investigate an approach that focuses on 
%generalizing 
%and
%composing 
identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency.

{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
re-use and extension of existing storage abstractions provided by the underlying software stack,
to enable the creation of new services via composition.
A programmable storage system can be realized by exposing existing functionality (such as 
metadata services and synchronization and monitoring capabilities) as interfaces that can be ``glued together''
in a variety of ways
using a high-level language. In contrast to programmable storage, 
\emph{active storage} differs through its focus on the injection and
execution of code within a storage system or storage device. Programmable 
storage differs from \emph{active storage}~\cite{riedel:vldb98} in that the 
former is applicable to any component of the storage system, while the latter 
focuses at the data access level (i.e. inject and execute arbitrary code at the 
I/O access layers). Given this contrast, we can say that active storage is an 
example of how one internal component (the storage layer) is exposed in a 
programmable storage system.

@@

Programmable storage systems can be viewed as an infrastructure for creating
abstractions to better separate policies from mechanisms. This idea is not new.
Software-defined networks (SDNs) create such an abstraction by separating the
control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
notion of control/data separation was also applied in software-defined storage
(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
filtering in OpenStack Swift. According to a SNIA white 
paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and 
facilitate flexible and dynamic provisioning of storage resources of different 
kinds, including flash memory and disk drives, to create a virtualized mapping 
between common storage abstractions (e.g. files, objects, and blocks) and 
storage devices taking data service objectives in terms of protection, 
availability, performance, and security into account. A programmable storage 
system exposes internal abstractions so that end users (not necessarily 
operators) can create new services on top of the storage stack. Thus, our notion 
of programmable storage differs from ``software-defined storage'' (SDS) in terms 
of goals and scope, although definitions of SDS are still in flux.
